---
title: "Capstone - Quiz3"
author: "Miguel Conde"
date: "Thursday, April 02, 2015"
output: html_document
---

```{r echo = FALSE, warning=FALSE, message=FALSE}
source("./R/perplex.R")
date()
```

Prepare:

```{r}
model <- loadModelTotal()

predict_init <- function() {
  setkey(model$m.1G, WG1)
  setkey(model$m.2G, WG1)
  setkey(model$m.3G, WG1, WG2)
  setkey(model$m.4G, WG1, WG2, WG3)
}

pp_init <- function() {
  setkey(model$m.1G, WG1)
  setkey(model$m.2G, WG1, WG2)
  setkey(model$m.3G, WG1, WG2, WG3)
  setkey(model$m.4G, WG1, WG2, WG3, WG4)
}

get_PPs <- function(phrase, w1, w2, w3, w4) {


  phrase1 <- paste(phrase, w1)
  phrase2 <- paste(phrase, w2)
  phrase3 <- paste(phrase, w3)
  phrase4 <- paste(phrase, w4)
  
  pp1 <- linePerplex4G(model, phrase1)
  pp2 <- linePerplex4G(model, phrase2)
  pp3 <- linePerplex4G(model, phrase3)
  pp4 <- linePerplex4G(model, phrase4)
  
  res1 <- data.frame(Words = c(w1, w2, w3, w4), PPs = c(pp1, pp2, pp3, pp4))
  
  pp1 <- linePerplex4G_v2(model, phrase1)
  pp2 <- linePerplex4G_v2(model, phrase2)
  pp3 <- linePerplex4G_v2(model, phrase3)
  pp4 <- linePerplex4G_v2(model, phrase4)
  
  res2 <- data.frame(Words = c(w1, w2, w3, w4), PPs_v2 = c(pp1, pp2, pp3, pp4))
  
  writeLines("PPs:\n")
  print(res1[order(res1$PPs), ])
  writeLines("\nPPs_v2:\n")
  print(res2[order(res2$PPs_v2), ])
}
```

Using model v7.5

First with **Total model**:

# Question 1

When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd

- die
- eat
- give
- sleep

Predict:

```{r Q1_1, echo =FALSE}
predict_init()
words <- c("live", "and", "I'd") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```


```{r Q1_2, echo = FALSE}
pp_init()

phrase <- "When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd"

get_PPs(phrase, "die", "eat", "give", "sleep")
```


Da "give", pero la letra de la canción dice "die"...

# Question 2

Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his

- spiritual
- marital
- financial
- horticultural

```{r Q2_1, echo =FALSE}
predict_init()
words <- c("me", "about", "his") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q2_2, echo = FALSE}
pp_init()

phrase <- "Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his"

get_PPs(phrase, "spiritual", "marital", "financial", "horticultural")
```

"financial"

# Question 3

I'd give anything to see arctic monkeys this

- morning
- weekend
- decade
- month

```{r Q3_1, echo =FALSE}
predict_init()
words <- c("arctic", "monkeys", "this") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q3_2, echo = FALSE}
pp_init()

phrase <- "I'd give anything to see arctic monkeys this"

get_PPs(phrase, "morning", "weekend", "decade", "month")
```

"morning", con muy poca dferenia sobre "weekend"

# Question 4

Talking to your mom has the same effect as a hug and helps reduce your

- sleepiness
- hunger
- stress
- happiness

```{r Q4_1, echo =FALSE}
predict_init()
words <- c("helps", "reduce", "your") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q4_2, echo = FALSE}
pp_init()

phrase <- "Talking to your mom has the same effect as a hug and helps reduce your"

get_PPs(phrase, "sleepiness", "hunger", "stress", "happiness")
```

Da "happiness" pero no puede ser; seguramente es "stress"

# Question 5

When you were in Holland you were like 1 inch away from me but you hadn't time to take a

- walk
- minute
- picture
- look

```{r Q5_1, echo =FALSE}
predict_init()
words <- c("to", "take", "a") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q5_2, echo = FALSE}
pp_init()

phrase <- "When you were in Holland you were like 1 inch away from me but you hadn't time to take a"

get_PPs(phrase, "walk", "minute", "picture", "look")
```

"look" por un pelo, sobre "picture"

# Question 6

I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the

- case
- matter
- account
- incident

```{r Q6_1, echo =FALSE}
predict_init()
words <- c("to", "settle", "the") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q6_2, echo = FALSE}
pp_init()

phrase <- "I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the"

get_PPs(phrase, "case", "matter", "account", "incident")
```

"case"

# Question 7

I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each
 
- toe
- arm
- hand
- finger

```{r Q7_1, echo =FALSE}
predict_init()
words <- c("groceries", "in", "each") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q7_2, echo = FALSE}
pp_init()

phrase <- "I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each"

get_PPs(phrase, "toe", "arm", "hand", "finger")
```

Empate virtual entre "arm" y "hand"; PP da lo mismo (105.7636) para ambos, PP_v2 prefiere "hand" (268.8135 vs 269.9055)

# Question 8

Every inch of you is perfect from the bottom to the

- center
- side
- middle
- top

```{r Q8_1, echo =FALSE}
predict_init()
words <- c("bottom", "to", "the") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

"top"

```{r Q8_2, echo = FALSE}
pp_init()

phrase <- "Every inch of you is perfect from the bottom to the"

get_PPs(phrase, "center", "side", "middle", "top")
```

# Question 9

I’m thankful my childhood was filled with imagination and bruises from playing

- weekly
- outside
- daily
- inside

```{r Q9_1, echo =FALSE}
predict_init()
words <- c("bruises", "from", "playing") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q9_2, echo = FALSE}
pp_init()

phrase <- "I’m thankful my childhood was filled with imagination and bruises from playing"

get_PPs(phrase, "weekly", "outside", "daily", "inside")
```

"outside"

# Question 10

I like how the same people are in almost all of Adam Sandler's

- stories
- pictures
- movies
- novels

```{r Q10_1, echo =FALSE}
predict_init()
words <- c("of", "Adam", "Sandler's") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q10_2, echo = FALSE}
pp_init()

phrase <- "I like how the same people are in almost all of Adam Sandler's"

get_PPs(phrase, "stories", "pictures", "movies", "novels")
```

Da "stories" por los pelos sobre "movies"

***

Let's try now with **Twitter model**:

```{r echo = FALSE}
rm(model)
gc()
model <- loadModelTwitter()
```

# Question 1

When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd

- die
- eat
- give
- sleep

Predict:

```{r Q1_1t, echo =FALSE}
predict_init()
words <- c("live", "and", "I'd") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```


```{r Q1_2t, echo = FALSE}
pp_init()

phrase <- "When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd"

get_PPs(phrase, "die", "eat", "give", "sleep")
```


Da "give", pero la letra de la canción dice "die"...

Twitter coincide

En principio, preferencia de Total: "die" - NOK

Cambiamos a "die" - OK

Total PP 0 - Total PP_v2 0/10 - Twitter PP 0/10 - Twitter PP_v2 0/10

# Question 2

Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his

- spiritual
- marital
- financial
- horticultural

```{r Q2_1t, echo =FALSE}
predict_init()
words <- c("me", "about", "his") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q2_2t, echo = FALSE}
pp_init()

phrase <- "Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his"

get_PPs(phrase, "spiritual", "marital", "financial", "horticultural")
```

"financial"

Twitter dice "marital", empatado con "hortocultural"

En principio, preferencia de Total: "financial" - NOK

??????????? A ojo, "marital" parece más apropiado... - OK

Total PP 0 - Total PP_v2 0/10 - Twitter PP 0.5/10 - Twitter PP_v2 0.5/10

# Question 3

I'd give anything to see arctic monkeys this

- morning
- weekend
- decade
- month

```{r Q3_1t, echo =FALSE}
predict_init()
words <- c("arctic", "monkeys", "this") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q3_2t, echo = FALSE}
pp_init()

phrase <- "I'd give anything to see arctic monkeys this"

get_PPs(phrase, "morning", "weekend", "decade", "month")
```

"morning", con muy poca diferenia sobre "weekend"

Twitter dice "weekend", casi empatado con "morning"

En principio, preferencia de Total: "morning" - NOK

Cambiamos a "weekend" - OK

Total PP 0 - Total PP_v2 0/10 - Twitter PP 1.5/10 - Twitter PP_v2 1.5/10

# Question 4

Talking to your mom has the same effect as a hug and helps reduce your

- sleepiness
- hunger
- stress
- happiness

```{r Q4_1t, echo =FALSE}
predict_init()
words <- c("helps", "reduce", "your") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q4_2t, echo = FALSE}
pp_init()

phrase <- "Talking to your mom has the same effect as a hug and helps reduce your"

get_PPs(phrase, "sleepiness", "hunger", "stress", "happiness")
```

Da "happiness" pero no puede ser; seguramente es "stress"

Twitter coincide

En principio, preferencia de Total: "happiness" - NOK

Estaba claro: cambiamos a "stress", aunque para Total y Twitter es la 3ª - OK

Total PP 0 - Total PP_v2 0/10 - Twitter PP 1.5/10 - Twitter PP_v2 1.5/10

# Question 5

When you were in Holland you were like 1 inch away from me but you hadn't time to take a

- walk
- minute
- picture
- look

```{r Q5_1t, echo =FALSE}
predict_init()
words <- c("to", "take", "a") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q5_2t, echo = FALSE}
pp_init()

phrase <- "When you were in Holland you were like 1 inch away from me but you hadn't time to take a"

get_PPs(phrase, "walk", "minute", "picture", "look")
```

"look" por un pelo, sobre "picture"

Twitter coincide

En principio, preferencia de Total: "look" - NOK

Pues cambiaremos a "picture" - OK

Total PP 0 - Total PP_v2 0/10 - Twitter PP 1.5/10 - Twitter PP_v2 1.5/10

# Question 6

I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the

- case
- matter
- account
- incident

```{r Q6_1t, echo =FALSE}
predict_init()
words <- c("to", "settle", "the") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q6_2t, echo = FALSE}
pp_init()

phrase <- "I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the"

get_PPs(phrase, "case", "matter", "account", "incident")
```

"case"

Twitter coincide

En principio, preferencia de Total: "case" - NOK

Pues iremos a la 2ª opción de todos: "matter" - OK

Total PP 0 - Total PP_v2 0/10 - Twitter PP 1.5/10 - Twitter PP_v2 1.5/10

# Question 7

I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each
 
- toe
- arm
- hand
- finger

```{r Q7_1t, echo =FALSE}
predict_init()
words <- c("groceries", "in", "each") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q7_2t, echo = FALSE}
pp_init()

phrase <- "I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each"

get_PPs(phrase, "toe", "arm", "hand", "finger")
```

Empate virtual entre "arm" y "hand"; PP da lo mismo (105.7636) para ambos, PP_v2 prefiere "hand" (268.8135 vs 269.9055)

Twitter prefiere "hand" claramente

En principio, preferencia de Total: como hay un empate y Twitter dice "hand": "hand" - OK

Total PP 0.5/10 - Total PP_v2 1/10 - Twitter PP 3.5/10 - Twitter PP_v2 3.5/10

# Question 8

Every inch of you is perfect from the bottom to the

- center
- side
- middle
- top

```{r Q8_1t, echo =FALSE}
predict_init()
words <- c("bottom", "to", "the") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```



```{r Q8_2t, echo = FALSE}
pp_init()

phrase <- "Every inch of you is perfect from the bottom to the"

get_PPs(phrase, "center", "side", "middle", "top")
```

"top"

Twitter coincide

En principio, preferencia de Total: "top" - OK

Total PP 1.5/10 - Total PP_v2 2/10 - Twitter PP 4.5/10 - Twitter PP_v2 4.5/10

# Question 9

I’m thankful my childhood was filled with imagination and bruises from playing

- weekly
- outside
- daily
- inside

```{r Q9_1t, echo =FALSE}
predict_init()
words <- c("bruises", "from", "playing") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q9_2t, echo = FALSE}
pp_init()

phrase <- "I’m thankful my childhood was filled with imagination and bruises from playing"

get_PPs(phrase, "weekly", "outside", "daily", "inside")
```

"outside"

Twitter coincide

En principio, preferencia de Total: "outside" - OK

Total PP 2.5/10 - Total PP_v2 3/10 - Twitter PP 5.5/10 - Twitter PP_v2 5.5/10

# Question 10

I like how the same people are in almost all of Adam Sandler's

- stories
- pictures
- movies
- novels

```{r Q10_1t, echo =FALSE}
predict_init()
words <- c("of", "Adam", "Sandler's") 
predictNG_v2(model, words, 20, finalSample = FALSE, fullRes=TRUE)
```

```{r Q10_2t, echo = FALSE}
pp_init()

phrase <- "I like how the same people are in almost all of Adam Sandler's"

get_PPs(phrase, "stories", "pictures", "movies", "novels")
```

Da "stories" por los pelos sobre "movies"

Twitter: PPs da "movies" claramente, pero PP_v2 da, por los pelos, "pictures" sobre "movies"

En principio, preferencia de Total: "stories" - NOK

Estaba claro, cambiamos a "movies" - OK

Total PP 2.5/10 - Total PP_v2 3/10 - Twitter PP 5.5/10 - Twitter PP_v2 5.5/10

```{r echo=FALSE}
date()
```

